\LARGE \bf\vspace{-10mm} Face Recognition: Too Bias, or Not Too Bias?

### 2020-02-01 sections/analysis.tex
updated supplemental citations

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
\include{tables/ethnicity_far}
\glsreset{tar}
\glsreset{far}
\end{table}

The gender-based \gls{det} curve shows that there are different performances for \gls{m} and \gls{f} at a fixed threshold (dashed-line). Similar effects exist for the other curves as well (lines omitted to declutter). Many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produced at different operating points (\ie \gls{fpr}) \gls{m} and \gls{f}, which is undesirable.  The difference in \gls{fpr} is approximately a factor of 2-- quite large indeed. If this kind of difference is common in industry, one would expect a factor of 3 more false positives to be reported based on one subgroup than another, which has a strong potential to be the cause of media articles reporting bias in 
\gls{fr}~\cite{england2019,snow2018}.

@@

\begin{figure}[t!] 
	\centering    
	\includegraphics[trim=0in 0.0in 0in 0in,clip, width=\linewidth] {images/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}, which scores plotted next to each. Humans tend to be more successful at recognizing their own subgroup, with few exceptions (bottom). }}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
		\vspace{-5mm}
\end{figure}

>>>>

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
 \vspace{-2mm}
\include{tables/ethnicity_far}
\glsreset{tar}
\glsreset{far}
 \vspace{-3mm}

\end{table}

The gender-based \gls{det} curve shows a difference in performances between \gls{m} and \gls{f} with a fixed threshold (dashed-line). Similar effects exist for the other curves as well (lines omitted to declutter). For many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produces different operating points (\ie \gls{fpr}), which is undesirable.  The difference in \gls{fpr} is approximately a factor of two-- quite large indeed. If this is the case in an industrial system, one would expect a difference in about double the false positives to be reported on based on subgroup alone. The potential ramifications of such a bias should not be overlooked, which it has not as of lately-- gaining interest of even main-stream media ~\cite{england2019,snow2018}.

@@

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=\linewidth] {images/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}, which scores plotted next to each. Humans tend to be more successful at recognizing their own subgroup, with few exceptions (bottom). }}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
		\vspace{-5mm}
\end{figure}

### 2020-02-01 supplemental.tex
updated supplemental citations

\section{Other \gls{cnn} models.}
Variations in optimal threshold are shown shown to generalize to other models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  ArcFace as well. For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).

\begin{figure}[h!]
    \centering
    \includegraphics[width=.9\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
    \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fnp} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone. Very similar results are found for ArcFace in Fig.~\ref{fig:detcurves}.
    }\label{fig:sdm-appendix-a}
\end{figure}

Sample faces per subgroup of the \gls{bfw} dataset are shown in Fig.~\ref{fig:montage:app}.
\begin{figure}[h!]
    \centering
    \includegraphics[width=.8\linewidth]{images/facemontage.pdf}
    \caption{\textbf{Sample of \gls{bfw}}. Each row depicts a different gender, \gls{f} (top) and \gls{m} (bottom). Columns are grouped by ethnicity (\ie \gls{a}, \gls{b}, \gls{i}, and \gls{w}, respectfully).}
    \label{fig:montage:app}
\end{figure}

>>>>

\section{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exists across different models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
\begin{figure}[h!]
\vspace{-2mm}
    \centering
    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
    \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fnp} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
    \vspace{-3mm}
\end{figure}

Sample faces per subgroup of the \gls{bfw} dataset are shown in Fig.~\ref{fig:montage:app}.
\begin{figure}[h!]
\vspace{-2mm}
    \centering
    \includegraphics[width=.8\linewidth]{images/facemontage.pdf}
    \caption{\textbf{Sample of \gls{bfw}}. Each row depicts a different gender, \gls{f} (top) and \gls{m} (bottom). Columns are grouped by ethnicity (\ie \gls{a}, \gls{b}, \gls{i}, and \gls{w}, respectfully).}
    \label{fig:montage:app}
    \vspace{-3mm}
\end{figure}

\begin{thebibliography}{widest entry}
 \bibitem[1]{Cao18} Cao, Qiong, Li Shen, Weidi Xie, Omkar M. Parkhi, and Andrew Zisserman. "Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face & Gesture Recognition.} 2018.
  \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun. "Deep residual learning for image recognition." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
 \bibitem[3]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. "Squeeze-and-excitation networks." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
\end{thebibliography}
